---
# system(" quarto preview README.qmd --to default ") # --no-watch-inputs --no-browse
# https://quarto.org/docs/publishing/quarto-pub.html
# https://quarto.org/docs/dashboards/layout.html#pages
# quarto render README.qmd --to html,gfm
# quarto render README.qmd -P alpha:0.2 -P ratio:0.3
# https://quarto.org/docs/computations/parameters.html#rendering
# quarto render README.qmd --execute-params params.yml
# quarto typst compile README.typ
# https://quarto.org/docs/publishing/github-pages.html#github-action
# quarto publish gh-pages README.qmd --no-prompt # --to index.html 
# quarto publish quarto-pub document.qmd
# quarto::quarto_render("README.qmd", output_format = "all")
# https://r-pkgs.org/other-markdown.html#sec-readme
# https://quarto.org/docs/computations/parameters.html#knitr
params:
  alpha: 0.1
  ratio: 0.1
  title: pipeline for robotweath.com post 
  description: targets, webR and dashboard 
title: _`r params$title`_
subtitle: param alpha = `r params$alpha`
description:  _`r params$description`_
format:
  dashboard:
    # theme: darkly
    orientation: rows
    # theme: default
    # expandable: true
    # scrolling: false
    #logo: images/penguins.png
    #nav-buttons: [linkedin, twitter, github]
    nav-buttons:
      - icon: github
        href: https://gitlab.com/johngavin/robotweath.com
      # - reddit
  # https://quarto.org/docs/output-formats/typst.html#overview
  # ~/Documents_/note_taking_atom/projects/rix/quarto_typst.md
  # Typst creates PDF output
  typst:
    # https://nrennie.rbind.io/blog/making-pdf-with-quarto-typst-latex/
    # quarto use template nrennie/PrettyPDF
    nav-buttons:
      - icon: github
        href: https://gitlab.com/johngavin/robotweath.com
    keep-typ: true
    toc-title: Contnts
    toc: true
    toc-depth: 2
    number-sections: false
    # section-numbering: 1.1.a
    columns: 2
  html:
    theme: darkly
  # https://quarto.org/docs/computations/r.html#output-formats
    # paging for row/col overflow
    # pages => add navbar in dashboard 
    # df-print: paged
    toc: false
    code-fold: true
    code-summary: "Show code"
    fontsize: "15pt"
    # section-numbering: 1.1.a
    number-sections: true
    # css: styles.css
#output: github_document # /html_document # NB: OLD Rmarkdown
#filters:
  # https://r-posts.com/add-shiny-in-quarto-blog-with-shinylive/
#  - webr
webr:
  packages:
  # CRAN packages installed when page opened
  - "ggplot2"
  - "ggiraph"
  - "dplyr" # dplyr::`%>%`
  # - "tidyverse" # avoid to speed up pg load
  # - "tidygraph"
  # - "igraph"
  # - "ggdag"
execute:
  echo: true
  error: true
  warning: true
# embed-resources: true
editor: source
# Quarto with R => need rmarkdown => knitr engine
# engine: knitr
# knitr: 
#   # also global Knitr options using opts_knit
#   # https://quarto.org/docs/computations/r.html#knitr-options
#   opts_chunk: 
#     # https://yihui.org/knitr/options/
#     code-fold: true
#     comment: "#>"
#     # # https://quarto.org/docs/computations/execution-options.html#figure-options
#     fig.path: "man/figures/README-"
#     out.width: "100%"
#     dev: "cairo_pdf"
#     R.options:
#       knitr.graphics.auto_pdf: true
# # https://quarto.org/docs/computations/python.html
# # jupyter: python3
# #
---


<!-- README.md is generated from README.qmd. Please edit that file -->
<!---
  enclose ALL chunks with {webr-r} - not {r}
  then play each in webpage sequentially
  else {r} runs locally only
  ~/Documents_GitHub/ibrokers/vignettes/webr_dag_golem_owl.qmd
--->

<!-- badges: start -->
<!-- badges: end -->

Some text
[in HTML.]{.content-visible when-format="html"}
[in typst]{.content-visible when-format="typ"}
[in PDF.]{.content-visible when-format="pdf"}
[in GH markdown]{.content-visible when-format="gmd"}

# README "`r params$title`"
```{r title}
params$title
```


## Installation

+ development version from [GitHub](https://github.com/)
``` 
devtools::install_github("<repo>/<package>")
```

## Example
+  solve a common problem
+ This statement requires further illumination[^1].

[^1]: Here lies the deeper exploration of the statement, shining light on the intricacies and supporting information that enrich the narrative.

# About

*Click for details*
https://metr.github.io/autonomy-evals-guide/

## Row {.flow}

::: card {.content-visible when-format="gmd"}

::: card-header
About card-header
:::

::: card-body
This dashboard presents ... card-body
:::


## Row {.flow}

### Not About hdr 3

::: {.content-visible when-format="html"}
<!-- 
For README.md only (when-format="gmd")
See below (when-format="html") for  README.html
-->

.gmd content only i.e. not in .html

:::

:::

# tabset em

::: {.content-visible when-format="html"}

## tabsett
test 

::: {.panel-tabset}

## Galaxy Exploration

This tab might contain an overview of your data exploration endeavors, perhaps some narrative on the galaxies you've charted.

## Star Analysis

Here, delve into the analysis of star compositions. This could be an excellent place for a visual plot or detailed statistical breakdown.

In this tab, present your findings on various planets. Interactive charts or data tables could be particularly illuminating here.

<!-- inside the structure from last chunk -->

## Nebulae Visualization

Visualizing the intricate structures of various nebulae encountered.

```{r, eval = FALSE}
#| label: ggplot
#| echo: FALSE
# Create a ggplot object
p <- ggplot(mtcars, aes(x = wt, y = mpg, tooltip = paste("Weight:", wt, "MPG:", mpg), data_id = row.names(mtcars))) +
  # NB: NOT geom_point
  geom_point_interactive()
# geom_bar_interactive() geom_line_interactive()
ggiraph::girafe(ggobj = p)
# library(ggplot2) ; example(ggplot)
# ggplot(nebulae_data, aes(x=position, y=light_intensity)) + geom_point()
```

## diagrams
+ https://medium.com/number-around-us/arming-your-data-spaceship-essential-quarto-tips-and-tricks-for-the-modern-explorer-ead0fa787adb
  + TODO: modals, content on the margin, parallax sections, and various embeddings
```{mermaid}
graph TD;
    A[Data Exploration] -->|Transform| B(Data Modeling);
    B --> C{Insight Discovery};
    C -->|Yes| D[Communicate Results];
    C -->|No| E[Iterate Further];
```

```{dot}
digraph G {
    A -> B;
    B -> C;
    C -> D;
    D -> A;
}
```

:::


# robotwealth.com

Goal: build `{targets}` pipeline of
robotwealth.com blog post.[^2]
[^2]: [https://robotwealth.com/how-to-model-features-as-expected-returns/]

## expected return models

./R/robotwealth.com/expected_returns_models.R

-   features \~ expected returns
    -   objective
        -   predict then capitalise on future returns
    -   intuitive optimisation and implementation
    -   direct comparison of features
    -   framework to add/eval old/new signals
-   common *risk* models
    -   e.g. covariance estimates
-   expected return & risk models
    -   direct comparison with trading costs
    -   optimisation techniques to trade-off
        -   return, risk, and turnover given constraints

## Installation

Install the development version of robotwealth.com from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
# devtools::install_github("JohnGavin/robotwealth.com")
```


## File structure

```{r tree}
#| code-fold: true
# output: asis
c('dplyr', 'stringr', 'fs') |>
  sapply(library, 
    character.only = TRUE, 
    quietly = TRUE) %>% 
  invisible()
excl <- "(figures)|(result)|(objects)|[.](md)|[.](htm)|[.](png)|(files)|(meta)|(old)"
incl <- "(A-Z)|(R)|(data)|(README)|[.](md)|[.](htm)|[.](yaml)|[.](yml)|[.](qmd)|[.](nix)|[.](R)"
incl |> 
  fs::dir_tree(regexp = ., invert = F, recurse = F) |> 
  head(20)
```

## File sizes

```{r dir_info}
"." %>% 
  dir_info(recurse = T) %>% 
  mutate(modify_date =
    as.Date(modification_time)) %>% 
  select(path, size, modify_date) %>% 
  filter(path %>% 
    str_detect(pattern = incl, negate = F)) %>% 
  arrange(desc(size))
```

## Pipelines

## `targets::tar_make`

```{r tar_config_set}
#| echo: false
#| code-fold: true

# library(robotwealth.com)
library(targets)
# targets setup
scr <- "exp_ret_mdl.R"
str <- "exp_ret_mdl_store"
targets::tar_config_set(script = scr, store = str)
```
Targets code: `{r} scr`.\
Targets store: `{r} str`.

```{r tar_make}
targets::tar_make(callr_function = NULL)
```

## Print data

```{r perps_all}
tar_read(perps_all)
```

## Plot graphs

```{r gg_eq_curve_1}
# TODO: link each plot to 
# https://robotwealth.com/how-to-model-features-as-expected-returns/
#| label: gg_eq_curve_1 sdfsdf
#| fig-cap: Equity crv !expr 'paste("1", "2", sep = "/")'
# | fig-cap: "Equity curve 1/2)"
#| fig-alt: "Equity curve 1:"
#| echo: false
tar_read(gg_eq_curve_1)
```
```{r gg_eq_curve_2}
#| fig-cap: "Equity curve 2/2)"
#| fig-alt: "Equity curve 2:"
tar_read(gg_eq_curve_2)
```
```{r gg_trd_buf_param}
#| fig-cap: "Trade buffer parameter"
#| fig-alt: "Trade buffer parameter:"
tar_read(gg_trd_buf_param)
```
```{r gg_turnover}
#| fig-cap: "Trade buffer parameter"
#| fig-alt: "Trade buffer parameter:"
tar_read(gg_turnover)
```
```{r gg_orig_yes_cost}
#| fig-cap: "Orig with costs"
#| fig-alt: "Orig with costs."
tar_read(gg_orig_yes_cost)
```
```{r gg_returns_plot}
#| fig-cap: "Returns"
#| fig-alt: "Returns plot."
tar_read(gg_returns_plot)
```
```{r gg_weights_plot}
#| fig-cap: "weights"
#| fig-alt: "weights plot."
tar_read(gg_weights_plot)
```
```{r gg_rets_wghts}
#| fig-cap: "Returns and weights"
#| fig-alt: "weights plot."
# FIXME: patchwork not wokring
#tar_read(gg_rets_wghts)
```

```{r sessionInfo}
#| eval: true
#| code-fold: true
#| echo: false
#| label: sessionInfo
quarto::quarto_version()
sessioninfo::session_info(include_base = TRUE)
```

```{bash echo foo, echo = FALSE, output = FALSE}
# FIXME: echo to render output panel (background job)?
echo "foo" 
uname -a # Linux kernel version
```

:::

::: {.content-visible when-format="html"}
<!-- 
For README.md only (when-format="html")
See below (when-format="gmd") for  README.html
-->

<!-- 

``` 
{r reticulate, eval = FALSE}
#| code-fold: true
#| eval: false
# if (!requireNamespace(reticulate)) 
#   install.packages("reticulate")
# library(reticulate)
# reticulate::py_config()
# reticulate::py_discover_config()
# # renv::use_python() https://blog.rstudio.com/2019/11/06/renv-project-environments-for-r/#Integration_with_Python
# reticulate::py_version()
# TODO: v nix install?
# py_install(c("polars"))
# py_run_string('import polars') # fails
# repl_python() ; import polars # fails
```

```
{python sys.version, eval = FALSE}
#| eval: false
# import sys
# print(sys.version)
```
-->

:::

# Notes

::: {.block fill="luma(230)" inset="8pt" radius="4pt"}

This is a block with gray background and slightly rounded corners.
```{r R.version.string}
#| code-fold: true
R.version.string
```


https://quarto.org/docs/output-formats/typst.html#typst-blocks

:::


```{r quarto cli, include = FALSE}
# https://quarto.org/docs/computations/r.html#chunk-options
#| eval: false
#| code-fold: true
#| label: quarto cli
#| echo: true
# system(" quarto preview README.qmd --to default ") # --no-watch-inputs --no-browse
# 
# NB: knitr::opts_chunk$set in yaml header above
# knitr::opts_chunk$set( # now in yaml above
#   collapse = TRUE, comment = "#>",
#   fig.path = "man/figures/README-",
#   out.width = "100%")
```
