---
description: DAG, Golem and Owl 
title: webR and dashboard
# format: html
format:
  dashboard:
    orientation: rows
    theme: default
    # expandable: true
    # scrolling: false
    nav-buttons:
     - reddit
     - icon: gitlab
       href: https://gitlab.com/Acme
output: 
  html_document:
    number_sections: true
execute:
  echo: true
  error: true
# embed-resources: true
editor: source
engine: knitr
knitr: 
  opts_chunk: 
    collapse: true
filters:
  - webr
# CRAN packages installed when page opened
webr:
  packages:
  # - "ggdag"
  - "ggplot2"
  - "dplyr" # dplyr::`%>%`
  # - "tidyverse" # dplyr::`%>%`
  # - "tidyr"
  - "purrr"
  - "tidygraph"
  - "igraph"
  - "ggdag"
---
<!---
  enclose ALL chunks with {webr-r} - not {r}
  then play each in webpage sequentially
  else {r} runs locally only
--->

```{webr-r}
# hidden code chunk
# install WebR-ready version from R-universe
# when the page loads => context setup
#| context: setup
#| label: WebR-ready r-universe.dev on page load
#| echo: false
# webr::install("igraph", repos = c("https://igraph.r-universe.dev", "https://cran.r-project.org"))
#webr::install("ggdag", repos = "https://r-causal.r-universe.dev")
# library(ggdag)
```

# preamble
## row 1 
### Col 1 {height="20%"}
:::{.callout-note}
If R-Universe package, install deps manually. e.g. add 
[`crayon`](https://github.com/r-lib/crayon), 
a dependency of 
[`collateral`](https://github.com/jimjam-slam/collateral), to yaml 
[`webr.packages`](https://docs.r-wasm.org/webr/latest/packages.html)
:::
### Col 1 
:::{.callout-tip}
Tap on textbox code then change it.
:::
### Col 1 
:::{.callout-note collapse="true"}
## plots
+ basic example with base R plotting
+ install ggplot2
```{webr-r}
#| label: base plot
plot(sin, -pi, 2 * pi)
```
:::


# DAG 

## Sidebar {.sidebar}
+ Richard McElreath's scientific framework
+ causal inference & model-based stats 
+ analyze effect of smoking on birth weight
  + mosaicData::babies dataset 
  + `tidyverse` and `ggdag` packages
```{webr-r}
# require(tidyverse) # already loaded in yaml?
R.version.string %>% paste(1, .) %>% print()
R.version.string %>% paste(2, .) 
```

```{r sidebar 2}
# this chunk runs locally not in webpage
require(tidyverse)
R.version.string %>% paste(3, .) %>% print()
R.version.string %>% paste(4, .) 
print(R.version.string) 
```

## DAG: A directed acyclic graph
### Row 1 {height="50%"}
#### Column 1
+ draw a causal diagram 
  + relationship between smoking and birth weight
  + any potential confounders

#### Column 2
+ causal assumptions 
+ structure of data-generating process
+ identify 
  + confounding variables
  + sources of bias
  + potential interventions

### Row 3 {height="75%"}
```{webr-r}
#| label: DAG
#| fig-cap: |
#|   In our model of the data science process, you start with data import
#| fig-alt: |
#|   A diagram displaying the data science cycle: Import -> Tidy -> Understand 
require(ggdag)
dagg <- dagify(
  bwt ~ smoke + age + parity + race,
  smoke ~ age + race,
  age ~ race,
  labels = c(bwt = "Birth weight",
             smoke = "Smoking status",
             age = "Mother's age",
             parity = "Parity",
             race = "Race")
)
dagg
# (dagitty::randomDAG(5, .5))
#(ggdag(dagg) )
#print(ggdag(dagg))
#plot(ggdag(dagg))
```


# Golem 

## Golem: A statistical model or estimator 
+ approximates the causal model and data
+ specified using formulas, priors, likelihoods, and algorithms
+ validated by checking its fit, predictions, and diagnostics


# Owl 

## Owl: analyze the golem
+ set of procedures and tools 
+ answer the scientific questions
  + posterior summaries, hypothesis tests, effect sizes, counterfactuals, and sensitivity analyses
  + simulating data from the statistical model 
    + to check its validity and robustness


```{webr-r}
# hidden code chunk
# installs WebR-ready version from R-universe
# when the page loads => context setup
# NB: Create R objects while loading in here!
#| context: setup
#| label: collateral r-universe.dev
webr::install("collateral", repos = c("https://jimjam-slam.r-universe.dev"))
```

:::{.callout-note}
If R-Universe package, install dependencies manually. e.g. add 
[`crayon`](https://github.com/r-lib/crayon), 
a dependency of 
[`collateral`](https://github.com/jimjam-slam/collateral), to yaml 
[`webr.packages`](https://docs.r-wasm.org/webr/latest/packages.html)
:::
:::{.callout-tip}
Tap on textbox code then change it.
:::

:::{.callout-note collapse="true"}
## base plot
+ basic example with base R plotting
+ install ggplot2
```{webr-r}
#| label: base plot
plot(sin, -pi, 2 * pi)
```
:::

# Top-level dash-page 

:::{.panel-tabset}

## 1. Operate on groups

```{webr-r}
#| label: operate-groups
mtcars |>
  mutate(
    qlog = map_peacefully(data, ~ log(.$wt)))
```

## 2. Add messages

```{webr-r}
#| label: add-messages
mtcars |>
  mutate(
    qlog = map_peacefully(data, ~ log(.$wt)),
    warn_msg = purrr::map_chr(qlog,
      \(x) paste(x$warnings, collapse = ", ")))
```

:::

# Notes
```{webr-r}
fit = lm(mpg ~ am, data = mtcars)
summary(fit)
```

```{r sessionInfo}
#| collapse: TRUE
#| echo: FALSE
# this chunk runs locally not in webpage
R.version.string %>% paste(3, .) %>% print()
R.version.string %>% paste(4, .) 
print(R.version.string) 
sessionInfo()
```
